---
title: "invert_cleaning2"
output: html_document
date: "2023-11-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Uploading libraries
```{r}
##Clear R's memory to start with clean slate
rm(list = ls())

##Import libraries
library(tidyverse)
library(stringr)
library(readxl)
library(here)
library(janitor)
library(vegan)

##Set working directory and import data
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
```

#1
#Uploading 2009-2014 data
```{r}
#Uploading and organizing 2009-2014 data
split <- read_excel(("/Users/inezmangino/Desktop/santa cruz lter/data/SCI_Resource_Data_2009_2014_raw.xlsx"), sheet ="Split", na=c("","."))
rough <- read_excel(("/Users/inezmangino/Desktop/santa cruz lter/data/SCI_Resource_Data_2009_2014_raw.xlsx"), sheet ="Rough", na=c("","."))

#Separating the split and rough tabs from the raw excel data
split1 <-split %>%
  dplyr::select(-Sorter) %>%
  mutate(Site=ifelse(Site=="Twin W","Twin West",Site)) %>%
  mutate_if(is.numeric, ~replace_na(., 0))%>%
  pivot_longer(cols=`Gammarid_<0.5`:Sponge, names_to="spe",values_to ="measurement") %>% #not really length, just the value from the cell which is number in this case
  mutate(source="Split") 
rough1 <-rough %>%
  dplyr::select(-Sorter) %>%
  mutate(Site=ifelse(Site=="Twin W","Twin West",Site)) %>%
  mutate_if(is.numeric, ~replace_na(., 0)) %>%
  pivot_longer(cols=`Pachy_0.0-9.9`:Sponge, names_to="spe",values_to ="measurement") %>%
  mutate(source="Rough")

all <- bind_rows(split1,rough1)

all1 <- all %>%
  separate(spe, into = c("TAXA", "size"), sep="_(?=[^_]+$)")%>%
  mutate(count = str_extract(TAXA, "[0-9]*$"), TAXA = gsub('[0-9]+', '', TAXA), TAXA=trimws(TAXA)) %>%
  mutate(year=substr(as.character(Date),1,4)) %>%
  mutate(day=substr(as.character(Date),9,10)) %>% 
  mutate(month=substr(as.character(Date),6,7)) %>% 
  rename_with(toupper) %>%
  dplyr::select(YEAR, MONTH, DAY, DATE, SITE, DEPTH, REP, TAXA, SIZE, MEASUREMENT, COUNT, FRACTION, SOURCE) 

all1 <-clean_names(all1)
all1$site <- tolower(all1$site)
all1$taxa <- tolower(all1$taxa)
all1$source <- tolower(all1$source)

#Cleaning taxa names for the "size" data frame
all1 <- all1 %>% 
  mutate(taxa=ifelse(taxa== "crab larvae", "crab", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "crabs", "crab", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "crabs count", "crab number", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "crab number", "crab", taxa)) %>%
  mutate(taxa=ifelse(taxa== "crabs number", "crab", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "urchin count", "urchin", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "urchins", "urchin", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "urchins number", "urchin", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "unknown isopod", "isopod", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "isopods", "isopod", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "isopods number", "isopod", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "gammarids", "gammarid", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "gammarids number", "gammarid", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "srs number", "srs", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "lrs number", "lrs", taxa))

#Creating separate count and size data frames because there were counts under the original measurement column.  
measured_taxa_rough <- c("crab",	"urchin", "srs", "lrs", "isopod", "gammarid")
measured_taxa_split <- c("crab", "urchin")

#Count data (w/o counts from measured species)
count_dt <- all1 %>% 
  filter(!((taxa %in% measured_taxa_rough & source=="rough")|(taxa %in% measured_taxa_split & source=="split"))) 
count_dt <- subset(count_dt, select= -count)
colnames(count_dt)[10] = "count"

#Count data (w/counts from measured species) 
count_dt2 <- all1
count_dt2 <- subset(count_dt2, select= -count)
colnames(count_dt2)[10] = "count"

#Size data
size_dt <- all1 %>% 
  filter((taxa %in% measured_taxa_rough & source=="rough")|(taxa %in% measured_taxa_split & source=="split"))
size_dt <- subset(size_dt, select = -size)
size_dt <-subset(size_dt, select = -count)
colnames(size_dt)[9] = "size"
```

#2a
#Uploading 1982-2008 count data
```{r}
#Uploading 1982-2008 count data
count_1982_2008 <- read_csv(here("data", "SCI_Scrapes_Counted_Invertebrates_1982-2009.csv"))

count_1982_2008 <- count_1982_2008 %>% 
   mutate(SITE=ifelse(SITE=="Twin W","Twin West", SITE)) %>% 
   mutate(TAXA = gsub('[].[]', ' ', TAXA), TAXA=trimws(TAXA)) %>% 
   mutate_if(is.numeric, ~replace_na(., 0)) %>% 
   mutate(month=substr(as.character(DATE),6,7)) %>%
   mutate(day=substr(as.character(DATE),9,10)) %>% 
   rename_with(toupper) %>%
   dplyr::select(YEAR,MONTH,DAY,DATE,SITE,DEPTH,REP,TAXA,SIZE_CLASS_MG,COUNT,FRACTION_SORTED,SOURCE) %>% 
   rename(size = SIZE_CLASS_MG)

count_1982_2008 <- clean_names(count_1982_2008)
count_1982_2008$site <- tolower(count_1982_2008$site)
count_1982_2008$taxa <- tolower(count_1982_2008$taxa)
count_1982_2008$source <- tolower(count_1982_2008$source)

#removing 2009 from this data frame because it was less complete than in the 2009-2014 file
count_1982_2008 <- subset(count_1982_2008, year != 2009)
colnames(count_1982_2008)[11] = "fraction"
```

#3a
#Combining all years of count data
```{r}
#Combining all years of count data
count_dt <- count_dt %>% 
  mutate(year=as.numeric(year)) %>% 
  mutate(count=as.numeric(count)) 

count_1982_2008 <- count_1982_2008 %>% 
  mutate(depth=as.numeric(depth)) %>% 
  mutate(rep=as.numeric(rep)) %>% 
  mutate(count=as.numeric(count)) 

count_all <- bind_rows(count_dt, count_1982_2008)
  
count_all <- count_all[!(count_all$taxa %in% "unknown"),]
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
count_dt2 <- count_dt2 %>% 
  mutate(year=as.numeric(year)) %>% 
  mutate(count=as.numeric(count)) 

count_1982_2008 <- count_1982_2008 %>% 
  mutate(depth=as.numeric(depth)) %>% 
  mutate(rep=as.numeric(rep)) %>% 
  mutate(count=as.numeric(count)) 

count_all2 <- bind_rows(count_dt2, count_1982_2008)
  
count_all2 <- count_all2[!(count_all2$taxa %in% "unknown"),]
```

#4a
#Cleaning taxa names and adding taxonomic info
```{r}
#Cleaning counted names and adding taxonomic info
 count_all <- count_all %>% 
 mutate(taxa=ifelse(taxa== "amphipoda", "amphipod", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "caprellids", "caprellid", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "sea anem", "sea anenome", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "anthuroid isopod", "anthuroid", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "bivalves", "bivalve", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "brittle stars", "brittle star", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "caprellids", "caprellid", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "chaet", "polychaete", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "poly", "polychaete", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "poly.", "polychaete", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "cirolanid isopod", "cirolanid", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "coleortera", "coleoptera", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "copepods", "copepod", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "crab larvae", "crab", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "crust", "crustacean", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "crusta", "crustacean", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "cuke", "cucumaria", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "cyphonautes", "bryozoan", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "fish larv.", "fish", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "gastro", "gastropod", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "gastro.", "gastropod", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "gastropods", "gastropod", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "idoteaids", "idotea", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "idoteid", "idotea", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "megalops", "crustacean", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "mysids", "mysid", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "ostracods", "ostracod", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "opisthob", "heterobranch", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "opistob", "heterobranch", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "pachy", "Pachythyone rubra", taxa)) %>%
  mutate(taxa=ifelse(taxa== "poly", "polychaete", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "poly.", "polychaete", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "sea anem", "anenome", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "sphaerm", "sphaeromatid", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "starfish", "sea star", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "tanaids", "tanaid", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "tube worm", "polychaete", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "unid worm", "unknown worm", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "unidentified", "unknown", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "unidentified amphipod", "amphipod", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "unidentified isopod", "isopod", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "unidentified worm", "unknown worm", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "unidentifiedworm", "unknown worm", taxa)) %>% 
  mutate(source=ifelse(source== "splits", "split", source)) %>% 
  mutate(taxa=ifelse(taxa== "jaeropsids", "jaeropsid", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "jaerop", "jaeropsid", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "sea anenome", "anenome", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "shl", "srs", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "shs", "srs", taxa)) 
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
 count_all2 <- count_all2 %>% 
  mutate(taxa=ifelse(taxa== "amphipoda", "amphipod", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "caprellids", "caprellid", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "sea anem", "sea anenome", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "anthuroid isopod", "anthuroid", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "bivalves", "bivalve", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "brittle stars", "brittle star", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "caprellids", "caprellid", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "chaet", "polychaete", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "poly", "polychaete", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "poly.", "polychaete", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "cirolanid isopod", "cirolanid", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "coleortera", "coleoptera", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "copepods", "copepod", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "crab larvae", "crab", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "crust", "crustacean", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "crusta", "crustacean", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "cuke", "cucumaria", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "cyphonautes", "bryozoan", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "fish larv.", "fish", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "gastro", "gastropod", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "gastro.", "gastropod", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "gastropods", "gastropod", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "idoteaids", "idotea", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "idoteid", "idotea", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "megalops", "crustacean", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "mysids", "mysid", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "ostracods", "ostracod", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "opisthob", "heterobranch", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "opistob", "heterobranch", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "pachy", "Pachythyone rubra", taxa)) %>%
  mutate(taxa=ifelse(taxa== "poly", "polychaete", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "poly.", "polychaete", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "sea anem", "anenome", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "sphaerm", "sphaeromatid", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "starfish", "sea star", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "tanaids", "tanaid", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "tube worm", "polychaete", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "unid worm", "unknown worm", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "unidentified", "unknown", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "unidentified amphipod", "amphipod", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "unidentified isopod", "isopod", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "unidentified worm", "unknown worm", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "unidentifiedworm", "unknown worm", taxa)) %>% 
  mutate(source=ifelse(source== "splits", "split", source)) %>% 
  mutate(site=ifelse(site== "frys i", "frys 1", site)) %>% 
  mutate(site=ifelse(site== "twin e", "twin east", site)) %>% 
  mutate(taxa=ifelse(taxa== "jaeropsids", "jaeropsid", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "jaerop", "jaeropsid", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "sea anenome", "anenome", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "shl", "srs", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "shs", "srs", taxa)) 

count_all2 <-subset(count_all2, select = -size)
```

#5a
#Sum together row with the same taxa name but different counts
```{r}
count_all2 <- count_all2[, c(1,2,3,4,5,6,7,8,10,11,9)]

count_all3 <- count_all2 %>% 
  group_by(year, month, day, date, site, depth, rep, taxa, fraction, source) %>% 
  mutate(across(c(count), sum)) %>% 
  distinct()
```

#6a
#Filtering out prey density sites/continuous dates for count data
```{r}
prey_density_sites <- c("bc 1", "diablo", "twin west")
continuous_years <- c("1995", "1996","1997", "1998", "1999", "2000", "2001", "2002", "2003", "2004", "2005", "2006", "2007", "2008", "2009", "2010", "2011", "2012", "2013", "2014")
oct_nov <- c("10", "11")

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
count_all_prey <- count_all3 %>% 
  filter(site %in% prey_density_sites)

count_all_prey_cont <- count_all_prey %>% 
  filter(year %in% continuous_years)

count_all3_cleaned <- count_all_prey_cont %>% 
  filter(month %in% oct_nov)
```

#2b
#Uploading 1982-2008 size data
```{r}
#Uploading 1982-2008 measurement data
measure_1982_2008 <- read_csv(here("data", "SCI_Scrapes_Measured_Invertebrates_1982-2009.csv"))

measure_1982_2008 <- measure_1982_2008 %>% 
   mutate(SITE=ifelse(SITE=="Twin W","Twin West", SITE)) %>% 
   mutate_if(is.numeric, ~replace_na(., 0)) %>% 
   mutate(month=substr(as.character(DATE),6,7)) %>%
   mutate(day=substr(as.character(DATE),9,10)) %>% 
   rename_with(toupper) %>%
   select(YEAR,MONTH,DAY,DATE,SITE,DEPTH,REP,TAXA,SIZE,COUNT,FRACTION_SORTED,SOURCE)

measure_1982_2008 <- clean_names(measure_1982_2008)
measure_1982_2008$site <- tolower(measure_1982_2008$site)
measure_1982_2008$taxa <- tolower(measure_1982_2008$taxa)
measure_1982_2008$source <- tolower(measure_1982_2008$source)

#removing 2009 from this data frame
measure_1982_2008 <- subset(measure_1982_2008, year != 2009)
colnames(measure_1982_2008)[11] = "fraction"
```

#3b
#Combining all years of size data
```{r}
size_dt <- size_dt %>% 
  mutate(year=as.numeric(year))

measure_1982_2008 <- measure_1982_2008 %>% 
  mutate(depth=as.numeric(depth)) %>% 
  mutate(rep=as.numeric(rep))

size_all <-bind_rows(size_dt, measure_1982_2008) 
size_all <-subset(size_all, select = -count)

# dropping 0's from this dataframe
size_all[size_all == 0] <- NA
size_all<- na.omit(size_all)
```

#4b
#Cleaning taxa names
```{r}
size_all <- size_all %>% 
  mutate(taxa=ifelse(taxa== "crabs", "crab", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "isopods", "isopod", taxa)) %>% 
  mutate(taxa=ifelse(taxa== "gammarids", "gammarid", taxa)) %>% 
  mutate(source=ifelse(source== "splits", "split", source)) %>% 
  mutate(site=ifelse(site== "frys i", "frys 1", site)) %>% 
  mutate(site=ifelse(site== "twin e", "twin east", site))
```

#6b
#Filtering out prey density sites/continuous dates for size data
```{r}
size_all_prey <- size_all %>% 
  filter(site %in% prey_density_sites)

size_all_prey_cont <- size_all_prey %>% 
  filter(year %in% continuous_years)

size_all_cleaned <- size_all_prey_cont %>% 
  filter(month %in% oct_nov)
```

#Adding in taxonomic information to count data
```{r}
names <- read_csv(here("data", "invertnamesall.csv"))

count_all3_cleaned2 <-count_all3_cleaned %>%
  left_join(names,by="taxa")
```

#EXPORT DATA
```{r}
write.csv(size_all_cleaned, "../size_all_cleaned_11282023.csv")
write.csv(count_all3_cleaned2, "../count_all3_cleaned_12082023.csv")
```

