---
title: "algae_explore"
output: html_document
date: "2023-08-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# uploading packages
```{r}
library(tidyverse)
library(stringr)
library(readxl)
library(here)
library(janitor)
library(vegan)
library(ggplot2)
library(dplyr)
library(lme4)
library(glmmTMB)
library(corrplot)
library(biostat3)
library(explore)
library(performance)
library(see)
library(patchwork)
library(car)
library(plotrix)
```

# uploading csv
```{r}
algae <- read_csv(here("data", "Algae_weight_All_Years_20230815.csv"))
```

# frequency summaries
```{r}
a1 <- main3_algae %>%
  distinct(year,site,depth,rep)%>%
  group_by(year,site, depth)%>%
  summarise(freq=n())

a2 <- main3_algae %>%
  distinct(year,site,depth)%>%
  group_by(year,site)%>%
  summarise(freq=n())

a3 <- main3_algae %>% 
  distinct(year, site) %>% 
  group_by(site) %>% 
  summarise(freq=n())

a4 <- algae %>% 
  distinct(year) %>% 
  group_by(year) %>% 
  summarise(freq=n())

a5 <- main3_algae %>% 
  distinct(year, rep) %>% 
  group_by(year) %>% 
  summarise(freq=n())

a6 <- main3_algae %>% 
  distinct(year, depth, rep) %>% 
  group_by(year, depth) %>% 
   summarise(freq=n())

a7 <- main3_algae %>% 
  distinct(year, site, rep) %>% 
  group_by(year, site) %>% 
  summarise(freq=n())

a8 <- main3_algae %>% 
  distinct(year, site, depth, rep) %>% 
  group_by(year, site, depth) %>% 
   summarise(freq=n())

# there are 23 options of species codes on each replicate of the RPC survey
```

# pulling out the three main prey density sites
# pulling out the most abundant algae species
## the other sites, dates, etc were randomly sampled
```{r}
unique(algae$site)

prey_density_sites <- c("bc 1", "diablo", "twin west")

most_abundant_algae <- c("Gelidium robustum", "Articulated Corralines", "Eisenia", "Chondracanthus spp.", "Gelidium nudifrons", "Rhodymeniales", "Codium spp.", "Prionitis", "Laminaria farlowii", "Egregia", "Laurencia spp.", "Gym")

algae_main3 <- algae %>% 
  filter(site %in% prey_density_sites)
```

# linear model assumption testing
```{r}
# 1. linearity 
plot(mod_1, 1)
plot(mod_2, 1)

# 2. independence of errors. acf plot of residuals checks for autocorrelation
plot(mod_1, 2)
plot(mod_2, 2)

acf(resid(mod_1))
acf(resid(mod_2))

# 3. homoscedasticity. residuals vs. fitted values scatter plot
plot(mod_1, 3)
plot(mod_2, 3)

# 4. normality of residuals. QQ plot
plot(mod_1, 4)
plot(mod_2, 4)

hist(resid(mod_1))
hist(resid(mod_2))

plot(mod_1, 5)
plot(mod_2, 5)
```

# running a linear model ? 
```{r}
mod_1 <- lm(wet_wt ~ depth, data= algae_main3)
mod_2 <- lm(wet_wt ~ year, data= algae_main3)

summary(mod_1)
summary(mod_2)

plot(wet_wt ~ depth, data= algae_main3, pch=16)
plot(wet_wt ~ year, data= algae_main3, pch=16)
```

# calculations for data analysis: biomass, mean, sd, se, variance
```{r}
algae_stats <- algae_main3 %>%  
  group_by(year, site, depth, common_name, major_group) %>%
  summarise(mean_weight = mean(wet_wt), 
            sd_weight = sd(wet_wt), 
            var_weight = var(wet_wt),
            se_weight = std.error(wet_wt),
            meansd_l= mean_weight - sd_weight,
            meansd_u = mean_weight + sd_weight)


algae_stats_most_abundant <- algae_stats %>% 
  filter(common_name %in% most_abundant_algae)

algae_stats <- mutate(algae_stats, mean_biomass = mean_weight * 0.1)
```

```{r}
hist(algae_stats[["mean_weight"]], freq = FALSE)
lines(density(algae_stats[["mean_weight"]]))
```

# #plotting MEAN algae weight over: time, depth, site
```{r}
ggplot(data = algae_stats, aes(x = year, y= mean_weight)) + geom_line()

ggplot(algae_stats, aes(x= year, y= mean_weight)) + geom_point() + labs(x= "year", y= "Algae Weight") + theme_bw()

ggplot(data = algae_stats, aes(x = depth, y= mean_weight)) + geom_line()

ggplot(data = algae_stats, aes(x = site, y= mean_weight)) + geom_line()

ggplot(data = algae_stats, aes(x = reorder(common_name, -mean_weight), y= mean_weight)) +
  geom_point() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 

ggplot(data = algae_stats, aes(x = reorder(common_name, -mean_weight), y= mean_weight)) +
  geom_line()+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 

ggplot(data = algae_stats, aes(x = year, y = mean_weight, group = common_name, colour = depth)) +
  geom_line() + theme(axis.text.x = element_text(angle = 45, size = 5)) +
  facet_wrap(facets = ~reorder(common_name, -mean_weight)) + theme_bw()

ggplot(data = algae_stats, aes(x = year, y = mean_weight, group = common_name, colour = common_name)) +
  geom_line() + theme(axis.text.x = element_text(angle = 45, size = 5)) +
  facet_wrap(facets = ~reorder(common_name, -mean_weight)) + theme_bw()


# filtered by only most 12 abundant algae spp.
ggplot(data = algae_stats_most_abundant, aes(x = year, y = mean_weight, group = common_name, colour = common_name)) +
  geom_line() + theme(axis.text.x = element_text(angle = 45, size = 5)) +
  facet_wrap(facets = ~reorder(common_name, -mean_weight)) + theme_bw()
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

pdf("plot1.pdf")
ggplot(algae_stats, aes(x=year, y=mean_weight, group= common_name, colour= common_name)) +
  geom_line() + xlab("mean wet weight") + facet_wrap(facets= ~reorder(common_name, -mean_weight))
dev.off()

ggplot(algae_stats, aes(x=year, y=mean_weight, group= common_name, colour= common_name)) +
  geom_line() + xlab("mean wet weight") 

pdf("plot1.pdf")
ggplot(algae_stats, aes(x=depth, y=mean_weight, group= common_name, colour= common_name)) +
  geom_line() + xlab("mean wet weight") + facet_wrap( ~ common_name)
dev.off()
```

# most abundant algae graphs
```{r}
ggplot(data = algae_stats_most_abundant, aes(x = year, y = mean_weight, group = common_name, colour = common_name)) +
  geom_line() + theme(axis.text.x = element_text(angle = 45, size = 5)) +
  facet_wrap(facets = ~reorder(common_name, -mean_weight)) + theme_bw()

ggplot(data = algae_stats_most_abundant, aes(x = year, y = mean_weight, group = common_name, colour = depth)) +
  geom_line() + theme(axis.text.x = element_text(angle = 45, size = 5)) +
  facet_wrap(facets = ~reorder(common_name, -mean_weight)) + theme_bw()

ggplot(data = algae_stats_most_abundant, aes(x = year, y = mean_weight, group = common_name, colour = depth)) +
  geom_point() + theme(axis.text.x = element_text(angle = 45, size = 5)) +
  facet_wrap(facets = ~reorder(common_name, -mean_weight)) + theme_bw()
```


# EDA: histograms, qq, etc
```{r}
ggplot(data = algae_stats, aes(x=mean_weight)) +
  geom_histogram(fill= "steelblue", color= "black") +
  ggtitle("Histogram of algae weight")

qq_algae <- ggplot(algae_stats, aes(sample = mean_weight)) + geom_qq() +stat_qq() + stat_qq_line() + theme_bw()
qq_algae
```

# shiny app to explore data
```{r}
algae_stats %>%  explore()
```

# builing generalized linear model's
```{r}
glm(formula = mean_weight ~ algae_stats + common_name + depth + site, family =  "gaussian", data = algae_avg)
```

```{r}
mod1 <- glm(mean_weight ~ common_name, data = algae_stats)

summary(mod1)
#lets break this table down
```

# finding my distribution
```{r}
```

# smith textbook notes on EDA
```{r}
str(algae_stats)
colSums(is.na(algae_stats))
boxplot(mean_weight ~ common_name,
        data= algae_stats)

boxplot(mean_weight ~ depth,
        data= algae_stats)

#normality and homogeneity of the dependent variable
## 1. the distribution of a continuous variable can be visualized by dividing the x-axis into "bins" 
## and counting the number of observations in each bin as a frequency polygon
ggplot() + geom_freqpoly(data = algae_stats, aes(mean_weight), bins = 7)

## 2. homogeneity of variance is an even distribution of covariate values around the
## mean and is an important assumption of a Gaussian GLM.
leveneTest(algae_stats$mean_weight,
            algae_stats$common_name,
            location= c("median"), trim.alpha = 0.25)

# what percentage of the response variable are zero's?
sum(algae_stats$mean_weight == 0,
    na.rm = TRUE) * 100 / nrow(algae_stats)

#multicollinearity among covariates
vif(lm(mean_weight ~ site + year,
       data= algae_stats))


xyplot(mean_weight ~ year | site, data = algae_stats,
 layout = c(2,2), 
 xlab = list(label = "weight", cex = 1.2),
 ylab = list(label = "year",cex = 1.2),
 strip = function(bg = 'white', ...)
 strip.default(bg = 'white', ...),
 scales = list(alternating = TRUE,
 x = list(relation = "free"),
 y = list(relation = "same")),
 panel = function(x,y){
 panel.grid(h = -1, v = 2)
 panel.points(x,y, col = 1,
pch = 16,
 cex = 1.2)
 panel.abline(lm(y~x),
col = 1,
lwd = 5)})
```

```{r}
write.csv(algae_stats_most_abundant,"/Users/inezmangino/Desktop/SCI_LTER/data/algae_stats_most_abundant.csv", row.names=FALSE)
```



